---
title: "NHANES Challenge"
format: html
editor: visual
---

### Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(tidyverse)
library(haven)
library(here)
library(ggplot2)
library(janitor)
```

### Read-in Datasets for this Analysis:

-   DR1TOT_J.xpt as 'diet_behavior'
-   2017-2018_Hg-Blood.XPT as 'blood_hg'
-   2017-2018_Hg-Urine.XPT as 'urine_hg'
-   2017-2018_Diabetes.XPT as 'diabetes'

```{r}
diet_behavior <- read_xpt(here("data/nhanes_data/DR1TOT_J.XPT"))
blood_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Blood.XPT"))
urine_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Urine.XPT"))
diabetes <- read_xpt(here("data/nhanes_data/2017-2018_Diabetes.XPT"))
demographics <- read_xpt(here("data/nhanes_data/2017-2018_Demographics.XPT"))
```

### Subset Read-in Datasets

Subset 'diet_behavior' as 'diet'

```{r}
diet <- select(diet_behavior, SEQN, DRD360, DRD370B, DRD370BQ, DRD370Q, DRD370QQ)
```

Subset 'diabetes' as 'tiid'

```{r}
tiid <- select(diabetes, SEQN, DIQ010)
```

Subset 'blood_hg' as 'bhg'

```{r}
bhg <- select(blood_hg, SEQN, LBXIHG, LBDIHGSI, LBXBGE, LBXBGM)
```

Subset "urine_hg' as 'uhg'

```{r}
uhg <- select(urine_hg, SEQN, URXUHG)
```

### Merge Subsets Into A Working Dataframe as 'df'

```{r}
df <- list(diet, tiid, bhg, uhg)

df <- df %>% reduce(full_join, by = 'SEQN')
```

1.  Filter Dataframe df for the following:

```{r}
# Assuming your dataframe is named `nhanes_data`
df <- df %>%
  # Filter out rows where DIQ010 or DRD360 are NA
  filter(!is.na(DIQ010), !is.na(DRD370B)) %>%
  # Keep only rows where DIQ010 and DRD360 are 1 or 2
  filter(DIQ010 %in% c(1, 2), DRD370B %in% c(1, 2)) %>%
  # Recode 1 to "Yes" and 2 to "No" for DIQ010 and DRD360
  mutate(
    DIQ010 = ifelse(DIQ010 == 1, "Yes", "No"),
    DRD370B = ifelse(DRD370B == 1, "Yes", "No")
  )
  
```

2.  Do some exploratory data analysis

```{r}
str(df)
```

```{r}
summary(df)
```

```{r}
unique(df$DRD370B)
```

```{r}
head(df)
```

2.  Determine the prior probability that someone has diabetes (DIQ010)

    -   P(B)

    ```{r}
    df %>% 
      tabyl(DIQ010) %>% 
      adorn_totals("row")

    priorprobability_diabetes <- .129
    ```

3.  Determine the prior probability that someone eats tuna (DRD370B)

    -   P(A)

    ```{r}
    df %>% 
      tabyl(DRD370B) %>% 
      adorn_totals("row")

    prior_probability_eats_tuna <- .361
    ```

4.  Determine the likelihood of eating tuna and having diabetes

    -   L(B\|A)

    ```{r}
    df %>% 
      tabyl(DIQ010, DRD370B) %>% 
      adorn_percentages("row") 
    ```

5.  Determine posterior probability via bayes theorem

    -   P(B\|A)

```{r}
likelihood_eatingtuna_diabetes <- 0.358
```

```{r}
(priorprobability_diabetes *
likelihood_eatingtuna_diabetes) /
prior_probability_eats_tuna
```

```{r}

contingency_table <- df %>%
  tabyl(DIQ010, DRD370B) %>% 
  adorn_totals(where = "row") %>%
  adorn_percentages("row")


contingency_table


P_A <- contingency_table['Yes', "Total"]

diabetes
P_B <- contingency_table["Total", 'Yes']   # P(B) - probability of eating tuna
P_A_given_B <- contingency_table['Yes', 'Yes']

P_B_given_A <- (P_A_given_B * P_B) / P_A
P_B_given_A


```

1.  Filter Datafrane df for the following:

```{r}
df <- df %>%
  filter(is.na(DIQ010), !is.na(DRD370B)) %>%
  filter(DIQ010 %in% c(1, 2), DRD370B %in% c(1, 2)) %>%
  
mutate(
  DIQ010 = ifelse(DIQ010 == 1, "Has Diabetes", "No Diabetes"),
  DRD370B = ifelse(DRD370B == 1, "Consumes Ahi", "No Ahi")
)
```

## Bayes Theorem

B \<- Has Diabetes

A \<- Consumes Ahi

#### P(B)

The probability that a person has diabetes.

```{r}
df %>%
  tabyl(DIQ010)
```

```{r}
probability_has_diabetes <- 0.13
```

#### L(B\|A) = P(A\|B)

```{r}
df %>%
  tabyl(DRD370B, DIQ010) %>%
  adorn_percentages("col")
```

```{r}
probability_consumes_ahi_given_diabetes <- 0.36
```

### P(A)

```{r}
df %>%
  tabyl(DRD370B)
```

```{r}
probability_consumes_ahi <- 0.36
```

### P(B\|A)

```{r}
(probability_has_diabetes * probability_consumes_ahi_given_diabetes
/ probability_consumes_ahi)
```

### Diabetes & Age

```{r}
demographics <- demographics %>%
  select(SEQN, RIDRETH1, RIAGENDR, RIDAGEYR)
```

```{r}
nhanes_data <- left_join(df, demographics, by =
                           "SEQN")
```

```{r}
nhanes_data <- nhanes_data %>%
  mutate(age_bracket = ifelse(RIDAGEYR > 40,
                              "old", "young"))
```

```{r}
ggplot(nhanes_data, aes(x = RIDAGEYR, fill = DIQ010)) + geom_density(alpha = 0.5)
```

### P(B\|A)

B \<- has diabetes

A \<- old, over 40

### P(A)

Probability old, over 40

```{r}
nhanes_data %>% 
  tabyl(age_bracket)
```

```{r}
probability_old <- 0.53
```

#### L(B\|A) = P(A\|B)

Probability is over 40 given they have diabetes

```{r}
nhanes_data %>%
  tabyl(age_bracket, DIQ010) %>%
  adorn_percentages("col")
```

```{r}
proability_old_given_diabetes <- 0.96
```

### P(B\|A)

```{r}
(probability_has_diabetes * proability_old_given_diabetes /probability_old)
```
