---
title: "naive bayes penguins"
format: html
editor: visual
---

### Libraries

```{r, message = FALSE}
library(tidyverse)
library(bayesrules)
library(janitor)
library(e1071)
```

### Story

We’ll start our naive Bayes classification with just a single penguin. Suppose an Antarctic researcher comes across a penguin that weighs less than 4200g with a 195mm-long flipper and 50mm-long bill. Our goal is to help this researcher identify the species of this penguin, Adelie, Chinstrap, or Gentoo.

#### Let's Plot!

Make a visualization that will help the researcher determine the species based on the characteristics described above

```{r}
data(penguins_bayes)
penguins <- penguins_bayes
```

```{r}
colnames(penguins)
```

```{r}
head(penguins)
```

The first thing that comes to mind when identifying a species is the creation of a decision tree, but this would require machine learning. So, I must think of a more simple visualization that completes a similar task.

```{r}
ggplot(data = penguins, aes(x = above_average_weight, fill = species)) +
  geom_bar() # 0 is no, 1 is yes
```

```{r}
ggplot(data = penguins, aes(bill_length_mm, flipper_length_mm, color = body_mass_g)) +
  geom_point() + 
  geom_smooth(method = lm) +
  facet_grid(~species)
```

Looking at the graph above, we can assume that it is a chinstrap penguin. Now we will check the certainty of this.

```{r}
penguins <- penguins %>%
  mutate(above_average_weight = if_else
(above_average_weight == 1, "yes", "no"))
```

Our Bayesian strategy

P(B\|A)

-   P(Chinstrap \| below average weight)

```{r}
penguins %>%
  tabyl(above_average_weight, species) %>%
  adorn_percentages("row")
```

There are more variables that should be considered, so we will look at Naive Bayes

### Naive Bayes Classification

```{r}
naive_model_hints <- naiveBayes(species ~ flipper_length_mm + bill_length_mm +
                                  above_average_weight, data = penguins)

# species is what we are looking for, the following variables are what we are considering, and then penguins is the dataset
```

```{r}
our_penguin <- data.frame(bill_length_mm = 50, flipper_length_mm = 195, above_average_weight = "no")
```

### Asking our model to create a prediction of what species our penguin is

```{r}
predict(naive_model_hints, newdata = our_penguin, type = "raw")
```

99% chance that it is a Chinstrap penguin

### Testing our model for accuracy with a confusion matrix

```{r}
penguins <- penguins %>%
  mutate(predicted_species = predict(naive_model_hints, newdata = .))
```

This just added a column to the penguins data set recording what species our model predicted

```{r}
head(penguins)
```

Now let's check accuracy

```{r}
penguins %>%
  tabyl(species, predicted_species) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

The model is best at predicting Adelie penguins, worst at predicting Chinstrap penguins
