---
title: "Test Results"
format: html
editor: visual
---

#### Load Libraries

```{r, message=FALSE}
library(tidyverse)
library(janitor)
library(vembedr)
```

#### The Story

You test positive for a rare disease that only effects 0.001 (One in one thousand people).

So you ask the doctor:

-   How certain is it that I have this disease?

    -   The test correctly identifies 99% of people that have the disease and only incorrectly identifies 1% of people that don't have the disease

What are the chances that you actually have this disease?

-   Some would say 99%, the accuracy of the test

    -   What does bayes say?

$$
P(B \mid A) = \frac{P(B) L(B \mid A)}{P(A)} 
$$

B \<- Has Disease

A \<- Positive test result

P(B\|A) - The probability of having the disease given a positive test result

#### Simulate the Data

```{r}

set.seed(70)  # For reproducibility

# Parameters
n_patients <- 10000  # Total population size
n_diseased <- 10     # Number of patients with the disease
sensitivity <- 0.99  # True positive rate (sensitivity)
false_positive_rate <- 0.01  # False positive rate

# Step 1: Create the DataFrame with patients
patients <- data.frame(
  patient_id = 1:n_patients,
  has_disease = c(rep(1, n_diseased), rep(0, n_patients - n_diseased))  # 10 with the disease, rest without
)

# Shuffle the DataFrame to randomize patient order
patients <- patients %>%
  sample_frac(size = 1)

# Step 2: Simulate the test results based on disease status
patients <- patients %>%
  mutate(
    # Test result is positive if the person has the disease and the test is sensitive,
    # or if they don't have the disease but it's a false positive
    test_result = case_when(
      has_disease == 1 & rbinom(n_patients, size = 1, prob = sensitivity) == 1 ~ "positive",
      has_disease == 0 & rbinom(n_patients, size = 1, prob = false_positive_rate) == 1 ~ "positive",
      TRUE ~ "negative"
    )
  )




```

#### Apply Bayes Theorem in Class

B \<- Has Disease

A \<- Positive test result

P(B\|A) - The probability of having the disease given a positive test result

### P(B)

The probability that a person has the disease

```{r}
patients %>%
  tabyl(has_disease)
```

```{r}
probability_patients_with_disease <- 0.001
```

### P(A)

The probability that the patient has a positive test result

```{r}
patients %>%
  tabyl(test_result)
```

```{r}
probability_positive_test_result <- 0.01
```

#### L(B\|A) = P(A\|B)

Probability has positive test result given patient has the disease

```{r}
patients %>%
  tabyl(test_result, has_disease) %>%
  adorn_percentages("col")
```

```{r}
probability_positive_test_result_given_has_disease <- 0.9
```

### P(B\|A)

```{r}
(probability_patients_with_disease * probability_positive_test_result_given_has_disease
/ probability_positive_test_result)
```

#### Video

```{r}
embed_url("https://www.youtube.com/watch?v=R13BD8qKeTg")
```

#### What about two positive test results?

$$
P(\text{have disease} \mid \text{positive second test}) = \frac{P(\text{have disease after first positive}) \cdot P(\text{positive second test} \mid \text{have disease})}{P(\text{positive second test})}
$$

After watching the video, I have a better understanding of how to answer about the two positive test results. So first we start with the first positive test, after the first test is positive, it changes the belief the first person has the disease. Then the new probability of the person having the disease after the first positive is now the prior for the second test. Then we would apply Bayes theorem again with the second test. Which is seen in the equation as the probability that a person has the disease given this second positive test result.\
